A parallelised and distributed approach to implementing Conway's Game of Life

### 2022-12-06 report/article_2.tex
fixing Figure 5

\subsection{On the topic of analysis 1}

@@

%------------------------------------------------
\section{A lonesome worker}
\subsection{Logic}
Our single-worker distributed implementation moved much of the processing to said worker. From the distributor, we sent a 
single RPC call to the worker, processed GoL turns on the worker, then sent the final state back. Results are shown in Fig. 5.
\begin{figure}
  \includegraphics[width=\linewidth]{para-v-dist.png}
  \caption{1 worker vs. 1 thread}
  \label{fig:chart5}
\end{figure}
\subsection{On the topic of Analysis 4}
The differences between the benchmark times is not significant when compared to multi-threaded work. We 
identified an average of 4.54\% time decrease, which is most likely attributed to how the single-threaded workload had 
go-routines for Alive Cells and Key presses, whereas the single worker was Step 1 of Distributed, being "barebones".

@@

\begin{enumerate}[noitemsep]
  \item Listen for worker and controller connections
  \item On connect, call the worker/controller back (bidirectional)
  \item Listen for GoL process request from controller
  \item Iterate, dividing and sending work to each worker.
  \item After getting all parts back together, apply changes to broker internal state and divide new state up
\end{enumerate}
%--------------
\begin{itemize}
  \item ADD BENCHMARKS HERE
\end{itemize}
%--------------
\subsection{Analysis 5}
%--------------
\begin{itemize}
  \item ADD ANALYSIS HERE
  \item explain trends, averages, maybe std dev?
  \item if low std dev then time not affected too much by worker count
\end{itemize}
%--------------
\section{Tolerating faults}
\subsection{Managing workers}
In our previous implementation, we never checked for worker disconnects. This leaves us in a vulnerable
position - if a worker were to be terminated at any point, then we would have one section of changes missing.
This would show itself with a static portion on the SDL window. There is also the case of adding more workers. Either
way, it means we need to keep the order of workers. We present Algorithm 3 to solve this.
\begin{algorithm}
  \caption{Worker-diff checker}
  \begin{algorithmic}
    \State Require $ids \subseteq \mathbb{N}$
    \State Require $oldIds \subseteq \mathbb{N}$ 
    \State Let $idMap \gets Map$
    \For{$id \in ids$}
      \State $idMap[id] \gets addr$
    \EndFor
  \For{$id \in oldIds$}
      \If{idMap[id] = nil}
      \State \Return False
      \EndIf
    \EndFor
    \State \Return True
  \end{algorithmic}
\end{algorithm}

What this algorithm achieves is detection of disconnects and reconnects.
Given that we have the loop invariant (in turn iterator) "no two workers will have the same worker id",
we can use this algorithm, and if we detect a change, recalibrate the workers to adjust for new/missing worker. 
One may ask, does a worker disconnect not throw errors when calling RPC methods? In this case, we utilise \texttt{client.Go},
which notifies the calling process of errors without crashing. So, if error, then trigger recalibration.

\section{Implementing Halo Exchange}

% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
TO BE IMPLEMENTED
%------------------------------------------------

\section{benchmarking and critical analysis of the Halo Exchange}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
TO BE IMPLEMENTED
%------------------------------------------------

@@

Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas. 
\begin{itemize}
\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
\item Curabitur feugiat
\item turpis sed auctor facilisis
\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
\item Fusce fermentum, mi sit amet euismod rutrum
\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
\item Pellentesque bibendum pretium aliquet
\end{itemize}
\blindtext % Dummy text

Text requiring further explanation\footnote{Example footnote}.

@@

\blindtext % Dummy text

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

\blindtext % Dummy text

@@

A statement requiring citation \cite{Figueredo:2009dg}.
\blindtext % Dummy text

\subsection{Thanks}

\blindtext % Dummy text

@@

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

>>>>

\subsection{On the topic of Analysis 1}

@@

%------------------------------------------------
\section{A lonesome worker}
\subsection{Logic}
Our single-worker distributed implementation moved much of the processing to said worker. From the distributor, we sent a 
single RPC call to the worker, processed GoL turns on the worker, then sent the final state back. Results are shown in Fig. 5.
\begin{figure}
  \includegraphics[width=\linewidth]{1v1.png}
  \caption{1 worker vs. 1 thread}
  \label{fig:chart5}
\end{figure}
\subsection{On the topic of Analysis 4}
The absurd differences we observed in the 1 worker vs. 1 thread (distributed versus parallel), is due 
to a number of things:
\begin{enumerate}
  \item Broker and Workers were positioned in US-east-1, meaning North Virginia, which is 7000km away.
  This oversight caused the massive difference in results.
  \item RPC calls for 1 worker may have caused overhead (extremely minor compared to above)
  \item Controller was not positioned in US-east-1, as we were not able to successfully run the controller
  program on the EC2 instance, due to issues with SDL.
\end{enumerate}
Due to this, the results are statistically worthless.

@@

\begin{enumerate}[noitemsep]
  \item Listen for worker and controller connections
  \item On connect, call the worker/controller back (bidirectional)
  \item Listen for GoL process request from controller
  \item Iterate, dividing and sending work to each worker.
  \item After getting all parts back together, apply changes to broker internal state and divide new state up
\end{enumerate}
%--------------
% \begin{figure}
%   \includegraphics[width=\linewidth]{multiworker.png}
%   \caption{Multi-worker performance}
%   \label{fig:chart6}
% \end{figure}
%--------------
\subsection{Analysis 5}
%--------------
\begin{itemize}
  \item ADD ANALYSIS HERE
  \item explain trends, averages, maybe std dev?
  \item if low std dev then time not affected too much by worker count
\end{itemize}
%--------------
\section{Tolerating faults}
In our previous implementation, we never checked for worker disconnects. This leaves us in a vulnerable
position - if a worker were to be terminated at any point, then we would have one section of changes missing.
This would show itself with a static portion on the SDL window. There is also the case of adding more workers. Either
way, it means we need to keep the order of workers. We present Algorithm 3 to solve this.
\begin{algorithm}
  \caption{Worker-diff checker}
  \begin{algorithmic}
    \State Require $ids \subseteq \mathbb{N}$
    \State Require $oldIds \subseteq \mathbb{N}$ 
    \State Let $idMap \gets Map$
    \For{$id \in ids$}
      \State $idMap[id] \gets addr$
    \EndFor
  \For{$id \in oldIds$}
      \If{idMap[id] = nil}
      \State \Return False
      \EndIf
    \EndFor
    \State \Return True
  \end{algorithmic}
\end{algorithm}

What this algorithm achieves is detection of disconnects and reconnects.
Given that we have the loop invariant (in turn iterator) "no two workers will have the same worker id",
we can use this algorithm. If we detect a change, recalibrate the workers to adjust for new/missing worker. 
One may ask, does a worker disconnect not throw errors when calling RPC methods? In this case, we utilise \texttt{client.Go},
which notifies the calling process of errors without crashing. So, if error, then trigger recalibration by resetting each worker's
internal state.

\section{Implementing Halo Exchange}
\subsection{How Jesus Christ came to be}
Our first attempt wasn't exactly the definition of halo exchange. It used halos distributed by broker to
each worker, who then used the halos to update their internal state. Now, our implementation transfers the turn iterator
logic to the workers. The sequence of events happens as follows:
\begin{enumerate}[noitemsep]
  \item Worker connects to broker, and is initialised with parameters of controller.
  \item On initialisation, worker will dial the top and bottom neighbours (if any), then await work.
  \item On work, send its halos, and evolve its own slice given the neighbouring halos
  \item Results are pushed to broker and loop
\end{enumerate}
This oversimplified view of halo exchange is represented by a significant amount of implementation code.
We present our results with Fig. 6.

\subsection{Analysis 6 - Judgement Day}
An obvious pattern we see is that the time taken is significantly reduced. We find that there is a negative
correlation between workers and time - increasing workers decreases the runtime of the program. We argue that
these results stems from the reduced overhead of needing to use the broker to transfer halos; decreased workload
from the broker having to manage turns actually becoming a broker seen typically in Publish-Subscriber models.

@@

@@

@@

\begin{itemize}
  \item Memory sharing implementation for channels - we were able to create an alternative to channels, but
  this ended up more or less the same performance as a regular channel, if not worse. A strong area of improvement
  would be to rewrite this, and substitute all the other channels used with our implementation
  \item Parallelising workers - we did not manage to implement this in time. Since all the code for multithreading
  was in our parallel code, we could have imported the code then implemented it inside our workers.
\end{itemize}
\subsection{Thanks}
\begin{itemize}
  \item Lecturers and TAs for their support
  \item Monster Energy\texttrademark
  \item My missing sanity
\end{itemize}

@@
